Parameterization of the responses of subarctic European vegetation to key environmental variables for ozone risk assessment

### 2022-06-01 VegOzone.tex
Lisa's comments

\section{Results}
\label{sec:results}
In this section, we first compute the average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different localizations to MM (Section~\ref{subsec:res_local}). We use the $\mathrm{DO_3SE}$ model to compute the ozone dose and projected biomass loss due to ozone for the two recent GSs 2018/19 for each species and localization (Section~\ref{subsec:do3se_results}). Finally, we put the results into a climatological perspective (Section~\ref{subsec:climatologies}).

\subsection{Parameter localization}
\label{subsec:res_local}

We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}. A higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.

We propose the following metric. From the long term of meteorological data, we select two distinct periods: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the average relative stomatal conductance $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$). Presumably, $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In addition, a small deviation between noon and morning and a low standard deviation indicates higher robustness of this localization to variability in growing conditions. From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the subarctic, cold, and MM $f_\mathrm{T}$ are shown marked by an alternating background color. Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively. The standard deviation is denoted with error bars.

Based on the above criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score. The differences between the localized parameterizations and the MM are smallest for coniferous trees. This indicates that the MM Boreal parameterization is well suited for the subarctic climate at Svanhovd. For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate. Deciduous trees display also a notable increase in mean stomatal conductance using our localizations.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_func_opt_mean}
  %\includegraphics[width=8.3cm]{fig08}
  \caption{Proposed metric to test local climate derived response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
  \label{fig:javis_func_opt_mean}
\end{figure}

@@

We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our localized temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.

The results are comprehensively displayed in Fig.~\ref{fig:pody_rel}. Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ biomass reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. The different background colors indicate the temperature response function localization. The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8). Circles indicate simulations with localized GS, canopy height, and leaf dimensions. With open symbols, the effect on $\mathrm{POD_y}$ taking SWP (drought) into account is displayed.

We identified in general that an increasing cold tolerance represented by a shift in the temperature response function (MM $<$ cold $<$ subarctic) leads to an increase in $\mathrm{POD_1}$ in both probed years compared to the MM parameterization. The more extreme the temperature response function localization the higher $\mathrm{POD_1}$. In 2018, however, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to cold in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures. An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years). Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees). In all other model experiments SWP was hardly of relevance. A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) reduces $\mathrm{POD_1}$. Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.

For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. In 2018, this range is smaller ($(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript. The aperture angle ($\angle(+,0,-)$) indicates the sensitivity visually. A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8. This indicates that drought conditions can reduce the ozone damage risk.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. The different localizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate. We find higher $\mathrm{POD_1}$ in 2019 than in 2018. The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM amounts to $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference due to $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.

@@

The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years and is of the same order of magnitude. By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization. This means that the ozone damage risk could be possibly underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details). Establishing dedicated subarctic parameterizations are a must to properly assess ozone damage risk in these regions.

@@

We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature parameterization localization (\emph{subarctic}) for each PFT. The estimates for SWP=off are listed in Table~\ref{tab:biomass_reduction}. We defined an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$. Superscript refers to PPFD0.8 and subscript to PPFD1.2. The reported range corresponds to the effect of taking SWP into account or not. As SWP reduces $\mathrm{POD_1}$, biomass reduction is decreased when taking SWP into account. The biomass reduction ranges between $2.5\,\unit{\%}$ (coniferous, 2018) and $17.4\,\unit{\%}$ (deciduous, 2019). For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
Coniferous trees are the least affected PFT and the magnitude of total biomass reduction is almost independent of the choice of the parameterization within given uncertainties.

>>>>

\section{Results}
\label{sec:results}
In this section, we first estimate the multiannual average stomatal conductance $\left<g_\mathrm{sto}\right>$ relative to $g_\mathrm{max}$ for each PFT from the long term meteorological data according to Eq.~(\ref{eq:stomatal}) and compare the different local climate derived parameterizations to MM (Section~\ref{subsec:res_local}). We also use the $\mathrm{DO_3SE}$ model to compute the ozone dose and predicted biomass loss due to ozone for the two recent GSs 2018 and 2019 for each species and set of local parameters (Section~\ref{subsec:do3se_results}). Finally, we put the results on ozone damage risk in 2018 and 2019 into a climatological perspective meaning how common these conditions have been in the past. To this end, we analyze both the past seasonality and the deviation from the norm of temperature, precipitation radiant energy, and ozone for these two years (Section~\ref{subsec:climatologies}).

\subsection{Local climate derived parameterization}
\label{subsec:res_local}

We presume that perennial vegetation will likely maximize carbon acquisition during the short subarctic growing season. Because the MM version of the $\mathrm{DO_3SE}$ model currently does not simulate net photosynthesis $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011}. A higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio, thus, indicates a higher growth potential in the local climate.

To compare the different parameterizations, we propose the following metric: From the selected set of meteorological data, we define two distinct periods in the diurnal cycle: noon ($11\,\unit{am}-1\,\unit{pm}$ local time) and morning ($5-9\,\unit{am}$ local time) and calculate the $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ ratio (Eq.~(\ref{eq:stomatal}) assuming $f_\mathrm{SWP}=1$). That this choice is justified is shown in Section~\ref{subsec:do3se_results}. Presumably, $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ around noon (the highest light intensity) is a good proxy for \chem{CO_2} uptake efficiency (growth). In addition, a small deviation between noon and morning and a low standard deviation indicates a lower location induced variability in growing conditions. From left to right in Fig.~\ref{fig:javis_func_opt_mean}, results using the \emph{subarctic}, \emph{cold}, and MM $f_\mathrm{T}$ are shown highlighted by an alternating background color (linen-white-linen). Within each of these groups $f_\mathrm{light}$ categories are indicated by PPFD0.8 and PPFD1.2, respectively. The standard deviation is denoted with error bars.

Based on the above criteria, we identify \emph{subarctic}--PPFD0.8 as the parameterization with the highest score. The differences between the localized parameterizations and the MM are smallest for coniferous trees. This indicates that the MM Boreal parameterization is well suited for the subarctic climate at Svanhovd. For perennial grassland, the differences compared to MM are largest suggesting that the MM Central European parameterization is not appropriate in a subarctic climate. Deciduous trees display also a notable increase in mean stomatal conductance using our local climate derived parameterizations.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{javis_func_opt_mean}
  %\includegraphics[width=8.3cm]{fig08}
  \caption{Proposed metric to test local climate derived parameterizations for response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous trees; (b) coniferous trees; (c) perennial grassland.}
  \label{fig:javis_func_opt_mean}
\end{figure}

@@

We modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per PLA for the three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using the MM parameterization and our climate derived local temperature (\emph{cold}, \emph{subarctic}) and light (PPFD0.8, PPFD1.2) parameterizations.

The results are displayed in Fig.~\ref{fig:pody_rel}. Critical levels for reduction in biomass \citep[deciduous forest $4\,\unit{\%}$, coniferous forest $2\,\unit{\%}$, and grasslands $10\,\unit{\%}$ reduction adopted from][]{ICP:MappingManual2017,ESPR:Hayes2021} are shown as horizontal dashed lines. The different background colors indicate the the local climate derived temperature response function. The light response function localizations are denoted by $+$ (PPFD1.2) and $-$ (PPFD0.8). Circles indicate simulations with localized GS, canopy height, and leaf dimensions. With open symbols, the effect on $\mathrm{POD_y}$ of any soil water stress is accounted for.

We identified that, in general, an increasing cold tolerance represented by a shift in the temperature response function (MM~$<$~cold~$<$~subarctic) leads to an increase in $\mathrm{POD_1}$ in both investigated years (2018 and 2019) compared to the MM parameterization. The more extreme the temperature response function localization the higher $\mathrm{POD_1}$. In 2018, coniferous trees and perennial grassland display a reduction in $\mathrm{POD_1}$ for the subarctic parameterization compared to the cold paramerization in responds to the prescribed lower $g_\mathrm{sto}$ at higher temperatures. This is not seen in 2019. An earlier start and overall longer GS in 2019 leads to higher $\mathrm{POD_1}$ compared to 2018 despite the occurrence of more episodes of elevated \chem{O_3} in 2018 than in 2019 (see Section~\ref{subsec:anomalies} for details on the particular features of these two years). Taking into account SWP, $\mathrm{POD_1}$ was only notably reduced in 2018 (for deciduous trees). In all other model experiments SWP was hardly of relevance. A larger aperture of stomata at low light intensities (PPFD0.8) causes an increase in $\mathrm{POD_1}$ compared to MM while a smaller aperture (PPFD1.2) is leading to reduced $\mathrm{POD_1}$. Due to the shape of $f_\mathrm{light}$, a symmetric variation can lead to an asymmetric response in $\mathrm{POD_1}$.

For deciduous trees (Fig.~\ref{fig:pody_rel}a), all model experiments exceed the CL by ($5-16\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). In 2019, the lowest $\mathrm{POD_1}$ estimate (MM, PPFD1.2) amounts to $10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and the highest (subarctic, PPFD0.8) $21\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. In 2018, this range is smaller $(12-18)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. The sensitivity to $f_\mathrm{light}$ is largest in the subarctic parameterization and amounts to $^{+1.6}_{-2.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.9}_{-2.7}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. Herein, the difference of PPFD0.8 compared to MM $f_\mathrm{light}$ is denoted as superscript and for PPFD1.2 as subscript. A low soil water potential caused a reduction in $\mathrm{POD_1}$ in 2018 in particular when using PPFD0.8. This indicates that water limiting conditions can reduce the ozone damage risk.

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL by $(0-10)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$. The different climate derived local parameterizations do not show much impact compared to MM indicating that the MM's Boreal Norway spruce parameterization is well-suited for a subarctic climate. We find higher $\mathrm{POD_1}$ in 2019 than in 2018. The maximum difference in $\mathrm{POD_1}$ determined between subarctic and cold parameterizations with respect to MM amounts to $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum difference due to $f_\mathrm{light}$ amounts to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019.

@@

The magnitude of all described effects on $\mathrm{POD_1}$ varies between PFTs as well as years and is of the same order of magnitude. By construction, $\mathrm{POD_1}$ from localized parmeterizations is always larger than estimates with the MM parameterization. This means that the risk of ozone damage could be underestimated especially in more normal years like 2019 (see Section~\ref{sec:stats} for details). Establishing dedicated subarctic parameterizations are most likely required to properly assess risk.

@@

We estimated the biomass reductions in accordance to the CLs \citep{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme climate derived local temperature parameterization (subarctic) for each PFT. The estimates, assuming no soil water stress, are given in Table~\ref{tab:biomass_reduction}. We defined an uncertainty range from the difference in biomass reduction caused by the variation of $f_\mathrm{light}$. Superscript refers to PPFD0.8 and subscript to PPFD1.2. Herein, the reported range indicated by dots corresponds to the limiting effect by soil water stress. Because soil water stress reduces $\mathrm{POD_1}$, biomass reduction is also reduced. The biomass reduction ranges between $2.5\,\unit{\%}$ (coniferous, 2018) and $17.4\,\unit{\%}$ (deciduous, 2019). For deciduous trees this amounts to a reduction by up to $55\,\unit{\%}$ compared to MM.
Coniferous trees are the least affected PFT suggesting the Boreal MM parameterization is appropriate for this PFT.
